### 2023-03-14 owen/01_eda.qmd
add owen initial ideas, commit this

---
title: "Chess EDA"
format: html
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)

```

## Data

- Confirm the time controls for the Lichess data?

What types of data?
How many games involve titles?

```{r}
#| warning: false

lichess_data <- readRDS("../rdata/lichess_pilot.RData")
```

```{r}
table(lichess_data$Variant)

head(sort(table(lichess_data$TimeControl), decreasing = TRUE))

## this looks to be seconds + incrememnt, makes the most sense 
## could just take common blitz formats, 3 mins and 5 mins,
## would be 50k games still

head(sort(table(lichess_data$Event), decreasing = TRUE))

lichess_data %>% filter(Variant == "Standard") %>% 
  group_by(Event) %>% 
  count() %>% 
  arrange(-n)

```

Get all games from a specific player in a specific format,
to see if `RatingDiff` tracks how
the rating changes.

```{r}
lichess_data %>% filter(TimeControl == "180+0") %>% 
  filter(Event == "Rated Blitz game") %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(peng_elo = ifelse(White == "penguingim1", WhiteElo, BlackElo)) %>% 
  mutate(peng_change = ifelse(White == "penguingim1", WhiteRatingDiff,
                              BlackRatingDiff)) %>% 
  select(UTCDate, UTCTime, peng_elo, peng_change) %>%
  mutate(peng_elo = as.numeric(peng_elo)) %>% 
  mutate(peng_change = as.numeric(peng_change)) %>% 
  mutate(ELO_change = peng_elo + peng_change) %>% 
  head(n = 10)

### so RatingDiff gives you the starting ELO for the next game

```

- Casual games don't change the rating, so do we want to remove them from 
the data entirely? Would likely have a different role in winner effects,
etc.

```{r}
## get simpler data

simple_data <- lichess_data %>% 
  filter(Event %in% c("Rated Bullet game", "Rated Blitz game"))

lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  group_by(TimeControl) %>% 
  count() %>% 
  arrange(-n)

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

# bullet_60

# head(sort(table(bullet_60$White), decreasing = TRUE), 20)

## have a few different players with good number of games
## get list of all players, fit some sort of model using all players
## who play enough games

head(sort(table(c(bullet_60$White, bullet_60$Black)), decreasing = TRUE), 25)

## top 25 players possibly enough then, consider all games they are involved
## in

table(bullet_60$Result)

```

## Multilevel Logistic Regression

Similar to @ding2021nhl, we wish to fit a multilevel logistic 
regression to see if we can identify individual 
effects for chess players in this data.

To do this, we will consider the current history of an individual player
as their win ratio in their past $n$ games, where we will investigate
the sensitivity of these results to different choices of $n$.

We will identify focal players as those who were used to query the API,
i.e the `Username` for each game.

```{r, past-n-results}

## given a players data, determine how many of the previous games they won
## do this for a single player first of all
## only need to do this for the potential focal players

curr_user <- "penguingim1"

curr_games <-

```

<!-- ## Streaks -->

<!-- Calculate the length of winning streaks for any given player, -->
<!-- look at how these vary across rankings maybe? -->

```{r}
#| include: false
# get_streaks <- function(games, player_name) {
#   player_games <- games %>% 
#     filter(White == player_name | Black == player_name) 
#   
#   winner <- player_games$
#   ## then get their streaks in here also
# }

peng_games <- bullet_60 %>% 
  filter(White == "penguingim1" | Black == "penguingim1")

peng_games

peng_games$w <- ifelse(peng_games$Username == peng_games$White, 1, 0)

# result column
peng_games$result <- ifelse(peng_games$w == 1 & peng_games$Result == "1-0" | 
                         peng_games$w == 0 & peng_games$Result == "0-1", 1, 0)

peng_games <- peng_games %>% 
  arrange(UTCDate, UTCTime)

peng_games$result

### then count the streaks in this
## win and loss streaks

wins <- which(peng_games$result == 1)

s <- split(wins, cumsum(c(TRUE, diff(wins) != 1)))
win_streaks <- map_dbl(s, length) %>% as.vector()

## now need to get the losing streaks
losses <- which(peng_games$result == 0)
s <- split(losses, cumsum(c(TRUE, diff(losses) != 1)))

loss_streaks <- map_dbl(s, length) %>% as.vector()

sum(loss_streaks) + sum(win_streaks)

nrow(peng_games)
```

<!-- For a given game, get the proportion of games won by the White player -->
<!-- in their last $n$ games, where we can vary $n$ and see how -->
<!-- changing it changes the results. -->

<!-- - Would you want to count wins as white and/or wins as black then? -->

<!-- - General winner effect -->

<!-- - Pick a player as "focal", look at whether they win or lose, -->
<!-- accounting for their colour, their game history, the ranking  -->
<!-- difference to their opponent, and possibly also the  -->
<!-- titles, which could have some sort of psychological impact on the outcome. -->

```{r}
#| include: false

### get past n results

prev_games <- function(player, hist_len) {

}

curr_user <- "penguingim1"

curr_data <- bullet_60 %>% 
  filter(White == curr_user | Black == curr_user)

curr_data$w <- ifelse(curr_data$White == curr_user, 1, 0)

## this doesn't account for draws, gives them equal weight to a loss

## one thing you could do is compare results after a draw with
## results after a win or a loss, see how they differ
## if not effect, and control for everything, then they should be the same

curr_data$result <- ifelse(curr_data$w == 1 & curr_data$Result == "1-0" | 
                         curr_data$w == 0 & curr_data$Result == "0-1", 1, 0)

curr_data %>% 
  select(White, Black, WhiteElo, BlackElo, w, result, UTCDate, UTCTime) %>% 
  arrange(UTCDate, UTCTime)

## then use this to get curr win history for curr_user for each game
```

<!-- ## Hot Hand Fallacy -->

<!-- - Potential different probabilities of success in subsequent games,  -->
<!-- i.e depending on who opponent is, black vs white, etc -->

<!-- - Evidence for some individual effects but -->
<!-- not sure if there is anything out there for -->
<!-- Chess already? -->

<!-- ## Some simple Models -->

<!-- - Take previous 10 games, calculate win proportion in those games -->

<!-- - Use this as a predictor for outcome -->

<!-- - Just predict win probability for White, to control for colour -->

<!-- ```{r} -->
<!-- ## logistic regression for win or less based on previous result -->
<!-- ## controlling for opponents ranking, your ranking, colour -->
<!-- ## previous game -->
<!-- ## time since previous game maybe? -->
<!-- ## -->
<!-- ``` -->

<!-- - The tennis stuff maybe and compare it? -->

<!-- - Does how the game ended make a difference? Don't seem to have this for lichess -->

<!-- - Games in tournaments vs rated, other factors like this -->

<!-- ### References -->

<!-- - The hot hand in the wild -->

<!-- - Surprised by the Hot Hand Fallacy? -->
<!-- A Truth in the Law of Small Numbers -->