### 2023-04-25 owen/Write Ups/02_more_eda.qmd
trying out some new things, fitting some new models

---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  html:
    toc: true
    toc-location: left
  # pdf
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## Identifying Problems

An issue was identified with the current implementation,
where the `previous_win_history` used, which was supposed to
be the proportion of the past $n$ games won compared to the
players overall win percentage, contained the result of the current game
(the outcome). When this is corrected the strong equal positive winner
effects are removed, and we now see a variable effect, 
with some players still displaying positive winner effects. However,
several of the players no longer show such an effect.

## Permuting Data

A quick sanity check on the existing model is to see if there are 
still strong winner and loser effects present when we permute the 
ordering of the events. This can be checked quite easily.

```{r clean lichess data}
#| warning: false

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))
  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
  as_tibble()

```

```{r permute results}

## need to do this inside the `get_hist` function

## permute the results of all games for a given player before then
## computing the current win proportion

get_hist_perm <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(perm_result = sample(focal_result)) %>% 
    mutate(perm_win_prop = c(cumsum(perm_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(perm_result, n = prev_n)))
  hist_games
}

perm_games <- map_dfr(top_players, get_hist_perm, bullet_60, prev_n = 10) %>% 
  as_tibble()
```

```{r}
tidy_games %>% 
  ggplot(aes(focal_win_prop)) + geom_histogram() +
  labs(title = "True Win Proportions")

perm_games %>% 
  ggplot(aes(perm_win_prop)) + geom_histogram() +
  labs(title = "Permuted Win Proportions")

## there is some sort of difference here, but not sure how big it is

```

Can also plot the current previous $n$ win proportion over time. 
Under the permuted model this should look more like random noise.

```{r}

perm_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, perm_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Permuted Games")

tidy_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, focal_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Original Data")

## these look more random anyway

```

Now we fit the stan model to the permuted data and investigate what
the effects are.

```{r get stan data permuted}

perm_init_data <- perm_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         perm_win_prop, elo_diff, perm_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(perm_win_prop, default = 0) - mean(perm_win_prop)) %>% 
  filter(perm_result != 0.5)

# perm_init_data

stan_data_ave <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

```

```{r fit stan model}
#| eval: false

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_ave <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_ave$save_object("../model_fits/perm_model_n10.RDS")
```

Then we want to compare the estimates from this permuted model to the true 
data.

```{r}

fit3_ave <- readRDS("../model_fits/perm_model_n10.RDS")
fit3_orig <- readRDS("../model_fits/init_model_n10.RDS")

# fit3_ave$summary()

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
  labs(title = "Permuted Data")

mcmc_hist(fit3_orig$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
  labs(title = "True Data")

```

```{r}
#| echo: true

## for the permuted data
prob_positive(fit3_ave)

## for the real data
prob_positive(fit3_orig)

```

Permuted data gives effects which are all zero on average, which is reasonable.
This would be expected as permuting the ordering of the games should
remove any possible winner effects. Similarly, the 
probability of positive effects is closer to 0.5, which
would be expected under a model with no effects.

Another way to confirm that having no history should lead to no 
winner effects is shown below.
We now replace the historic win proportion with uniform 
draws on the interval $[-1, 1]$. The estimated winner effects 
under such a history are shown below.

```{r}
#| eval: false
stan_data_rand <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

random_win_prop <- runif(n = stan_data_rand$N, min = -1)

stan_data_rand$win_prop <- random_win_prop

fit3_rand <- mod$sample(data = stan_data_rand,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_rand$save_object(file = "../model_fits/unif_hist.RDS")

```

We also see that the probability of these effects being positive is
close to 0.5, as would be expected under this model.

```{r}

fit3_rand <- readRDS("../model_fits/unif_hist.RDS")

mcmc_hist(fit3_rand$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

# fit3_rand$summary()

prob_positive(fit3_rand)

```

## Separating out history

We want to next look at the history as both the current win 
streak, up to but not including the previous game, along with the result 
of the previous game, to see if we can identify how potential winner 
effects can be decomposed between these two components.

We do this with a model of the form

$$
P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + 
\delta_j x_{ij}^{*} + 
\gamma z_{ij}))},
$$

where now $x_{ij}^{*}$ is the result of the __previous__ game
played by player $j$, and $x_{ij}$ is their normalised 
win ratio over the previous $n$ games, not including the single
previous game.

```{r get data for winner effects}

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

```

```{r fit the stan model}
#| eval: false

stan_file <- "../model4.stan"

mod <- cmdstan_model(stan_file)

fit4 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit4$summary()

fit4$save_object(file = "../model_fits/model4.RDS")

```

```{r load the stan fit}

fit4 <- readRDS("../model_fits/model4.RDS")

mcmc_hist(fit4$draws(c("delta"))) +
  labs(title = "Effect of Previous Game")

mcmc_hist(fit4$draws(c("beta"))) +
  labs(title = "Effect of Previous n games, excluding most recent game")

```

This model seems to fit well, with no potential concerns also.

We can then again compare these estimated effects. This

```{r check whether positive effects}
#| echo: true

## the beta parameters (previous n, not last game)
prob_positive(fit4)

## the delta parameters (single game) 
prob_positive(fit4, param = "delta")
```

This seems to indicate that the previous game on its own does not have as clear
an effect, but that the streak of previous games shows stronger evidence
for a winner effect.

## Consider all bullet games

Previously we considered only 60 second bullet games. We wish to 
consider if there is evidence for this effect across all bullet games.

```{r get all bullet games}

all_bullet <- lichess_data %>% 
  filter(Event == "Rated Bullet game")

tidy_games <- map_dfr(top_players, get_hist, all_bullet, prev_n = 10) %>% 
  as_tibble()

tidy_games

stan_data <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
  filter(focal_result != 0.5)

```

```{r refit model 3 for all bullet}
#| eval: false

stan_data_ave <- list(N = nrow(stan_data),
                  J = length(top_players),
                  y = stan_data$focal_result,
                  id = stan_data$focal_id,
                  colour = stan_data$focal_white,
                  elo = stan_data$elo_diff,
                  win_prop = stan_data$ave_prop)

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_bullet <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_bullet$save_object(file = "../model_fits/model3_n10_all_bullet.RDS")
```

```{r}
fit3_bullet <- readRDS("../model_fits/model3_n10_all_bullet.RDS")

mcmc_hist(fit3_bullet$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

prob_positive(fit3_bullet)
```

How do these estimates compare to those with only the 60 second games?
The model with all bullet games has all posterior means greater than 0.
There is not much of an overall change between the estimated mean
effect sizes between the two datasets.
Using all bullet games also leads to 
smaller posterior standard deviations for the
winner effect size for all players.

```{r refit model 4 for all bullet games}
#| eval: false

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

stan_file <- "../model4.stan"

mod <- cmdstan_model(stan_file)

fit4 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit4$summary()

fit4$save_object(file = "../model_fits/model4_n10_all_bullet.RDS")

```

```{r examine this model fit}
fit4 <- readRDS("../model_fits/model4_n10_all_bullet.RDS")

mcmc_hist(fit4$draws(variables = c("beta", "mu1", "tau1", "gamma1", "gamma2")))

mcmc_hist(fit4$draws(variables = c("delta", "mu3", "tau3")))

prob_positive(fit4)
prob_positive(fit4, param = "delta")
```

## Just previous game

We could also only use the previous game as a predictor, 
without using the rest of the history.

```{r}
#| eval: false

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  prev_game = hist_data_init$prev_game)

stan_file <- "../model5.stan"

mod <- cmdstan_model(stan_file)

fit5 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit5$save_object(file = "../model_fits/model5_n10_all_blitz.RDS")

```

```{r}
fit5 <- readRDS("../model_fits/model5_n10_all_blitz.RDS")

mcmc_hist(fit5$draws(variables = c("delta")))

prob_positive(fit5, param = "delta")

```

Fitting using just the previous game as a predictor seems to give a reasonable
model, with a similar interpretation as when we include the history also.

For this model, a posterior mean value for delta of 0.25 corresponds to 7%
increase in the win probability having won the previous game compared 
to having lost it.

Not clear if this difference is due to excluding the win
history or due to using a larger/different dataset.

### Next Steps

- Can we compare the performance of these models, using loo-cv 
or similar?

- Could this model be parameterized better?

- Any other possible problems with this?

- What about some sort of mixture of the previous game and the others before 
that?

 - Some sort of decaying effect?