### 2023-04-28 owen/Write Ups/02_more_eda.qmd
updating the models and a bit more writing up on it also

---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  html:
    toc: true
    toc-location: left
  # pdf
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## Identifying Problems

An issue was identified with the current implementation,
where the `previous_win_history` used, which was supposed to
be the proportion of the past $n$ games won compared to the
players overall win percentage, contained the result of the current game
(the outcome). When this is corrected the strong equal positive winner
effects are removed, and we now see a variable effect, 
with some players still displaying positive winner effects. However,
several of the players no longer show such an effect.

### Permuting Data

A quick sanity check on the existing model is to see if there are 
still strong winner and loser effects present when we permute the 
ordering of the events. This can be checked quite easily.

```{r clean lichess data}
#| warning: false

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0")

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))
  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
  as_tibble()

```

```{r permute results}

## need to do this inside the `get_hist` function

## permute the results of all games for a given player before then
## computing the current win proportion

get_hist_perm <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(perm_result = sample(focal_result)) %>% 
    mutate(perm_win_prop = c(cumsum(perm_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(perm_result, n = prev_n)))
  hist_games
}

perm_games <- map_dfr(top_players, get_hist_perm, bullet_60, prev_n = 10) %>% 
  as_tibble()
```

```{r}
tidy_games %>% 
  ggplot(aes(focal_win_prop)) + geom_histogram() +
  labs(title = "True Win Proportions")

perm_games %>% 
  ggplot(aes(perm_win_prop)) + geom_histogram() +
  labs(title = "Permuted Win Proportions")

## there is some sort of difference here, but not sure how big it is

```

Can also plot the current previous $n$ win proportion over time. 
Under the permuted model this should look more like random noise.

```{r}

perm_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, perm_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Permuted Games")

tidy_games %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  group_by(focal_user) %>% 
  mutate(index = row_number()) %>% 
  ggplot(aes(index, focal_win_prop, colour = focal_user)) +
  geom_line() +
  facet_wrap(~focal_user, scales = "free_x") +
  labs(title = "Original Data")

## these look more random anyway

```

Now we fit the stan model to the permuted data and investigate what
the effects are.

```{r get stan data permuted}

perm_init_data <- perm_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         perm_win_prop, elo_diff, perm_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(perm_win_prop, default = 0) - mean(perm_win_prop)) %>% 
  filter(perm_result != 0.5)

# perm_init_data

stan_data_ave <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

```

```{r fit stan model}
#| eval: false

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_ave <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_ave$save_object("../model_fits/perm_model_n10.RDS")
```

Then we want to compare the estimates from this permuted model to the true 
data.

```{r}

fit3_ave <- readRDS("../model_fits/perm_model_n10.RDS")
fit3_orig <- readRDS("../model_fits/init_model_n10.RDS")

# fit3_ave$summary()

mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
  labs(title = "Permuted Data")

mcmc_hist(fit3_orig$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2"))) +
  labs(title = "True Data")

```

```{r}
#| echo: true

## for the permuted data
prob_positive(fit3_ave)

## for the real data
prob_positive(fit3_orig)

```

Permuted data gives effects which are all zero on average, which is reasonable.
This would be expected as permuting the ordering of the games should
remove any possible winner effects. Similarly, the 
probability of positive effects is closer to 0.5, which
would be expected under a model with no effects.

Another way to confirm that having no history should lead to no 
winner effects is shown below.
We now replace the historic win proportion with uniform 
draws on the interval $[-1, 1]$. The estimated winner effects 
under such a history are shown below.

```{r}
#| eval: false
stan_data_rand <- list(N = nrow(perm_init_data),
                  J = length(top_players),
                  y = perm_init_data$perm_result,
                  id = perm_init_data$focal_id,
                  colour = perm_init_data$focal_white,
                  elo = perm_init_data$elo_diff,
                  win_prop = perm_init_data$ave_prop)

random_win_prop <- runif(n = stan_data_rand$N, min = -1)

stan_data_rand$win_prop <- random_win_prop

fit3_rand <- mod$sample(data = stan_data_rand,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_rand$save_object(file = "../model_fits/unif_hist.RDS")

```

We also see that the probability of these effects being positive is
close to 0.5, as would be expected under this model.

```{r}

fit3_rand <- readRDS("../model_fits/unif_hist.RDS")

mcmc_hist(fit3_rand$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

# fit3_rand$summary()

prob_positive(fit3_rand)

```

## Initial Model

For completeness, we include the initial model output here again, to 
show the results.

$$
P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
$$

where:

- $\alpha_j$ is a player level random effect
- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
- $\gamma$ is a fixed effect of game level coefficients

We then partially pool the $\alpha$ and $\beta$ coefficients, so

$$
\alpha_j \sim \mathcal{N}(\mu_2, \tau_2)
$$
$$
\mu_2 \sim \mathcal{N}(0, 5)
$$
$$
\tau_2 \sim \mathcal{Cauchy}(0, 5)
$$

and similarly

$$
\beta_j \sim \mathcal{N}(\mu_1, \tau_1)
$$
$$
\mu_1 \sim \mathcal{N}(0, 5)
$$
$$
\tau_1 \sim \mathcal{Cauchy}(0, 5)
$$

We rescale the win ratio $x_{ij}$, by the average win ration of
player $j$. This means $\beta_j$ captures how a players win
probability changes as their historic performance deviates from 
the overall average win ratio.

For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
the focal player and the ELO difference between the focal 
player and their opponent. So that means we have 
$\gamma_1$, which tells us how the win probability changes in going
from black to white,
and $\gamma_2$ gives how the win probability changes as the focal player
has a better ELO than their opponent.

We show the posterior estimates for these parameters below.

```{r load in initial model here}

fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")

# fit3_ave$summary()

mcmc_hist(fit3_ave$draws(c("alpha", "mu2", "tau2")))
mcmc_hist(fit3_ave$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))
```

#### Interpreting this output

Given this output, we can get the probability a player who is black
will win against an equally ranked player, when their current win ratio 
is exactly their average will be $\frac{1}{1+\exp(-\alpha_j)}$.
So a positive $\alpha_j$ indicates a player favoured to win 
even when playing black. The overall $\mu_2$ tells us about the 
global probability of winning as black against an equally ranked
opponent, which is just below 0, indicating a slightly less than 50% 
chance of winning.
If a player
is playing as white their win probability will be
$\frac{1}{1+\exp(-(\alpha_j + \gamma_1))}$,
so $\gamma_1$ being positive indicates the win
probability increases when playing white, which makes sense.
Playing white makes a player who has 50% chance of winning as black
have approximately a 54% chance of winning with white. Wikipedia says the increase 
in win probability from playing white has been estimated to be 
4-6% so quite similar.

Similarly, $\gamma_2$ tells us about how the ELO of the focal player 
being larger than their opponent influences their win probability.
The estimate here indicates that a large ELO difference makes a player very
likely to win, as would be expected. For example, a player who
has a 50% chance of winning against an opponent with the same ELO
will have about a 61% chance of winning against an opponent with an ELO score
100 points lower.

#### How big are these effects

We then want to quantify how much these estimated effects change the 
expected win probability.
We have seen the size of the effects above for the non history parameters.
The largest posterior mean for a $\beta\approx 1.56$. Since the normalized win
proportion is between -0.6 and 0.6, that means that for a game
where this player has a 50% chance of winning with a normalized average win
proportion of 0, if their normalized win proportion is 0.5, the
win probability will increase by $\approx 19\%$.
This would occur, for example, if they win 50% of their total games but have
won all 10 of their last 10. This is a similar size effect to the 
increase in win probability which would come from playing an 
opponent with an ELO ranking that is 170 points lower.

Similarly, a normalized win history 0.25 
above average corresponds to a difference 
of approximately 80 in ELO scores.

The underlying parameter $\mu_1$ has posterior mean of 0.606. 
A normalised win history of 0.25 with a winner effect of this size
would increase the win probability by about 4%, a similar effect
as playing as white instead of black, all other covariates being
equal.

## Separating out history

We want to next look at the history as both the current win 
streak, up to but not including the previous game, along with the result 
of the previous game, to see if we can identify how potential winner 
effects can be decomposed between these two components.

We do this with a model of the form

$$
P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + 
\delta_j x_{ij}^{*} + 
\gamma z_{ij}))},
$$

where now $x_{ij}^{*}$ is the result of the __previous__ game
played by player $j$, and $x_{ij}$ is their normalised 
win ratio over the previous $n$ games, not including the single
previous game. Note that $x_{ij}^{*}$ is not normalised,
so can take on values $0,0.5,1$.

```{r get data for winner effects}

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

```

```{r fit model 4}
#| eval: false

stan_file <- "../model4.stan"

mod <- cmdstan_model(stan_file)

fit4 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit4$summary()

fit4$save_object(file = "../model_fits/model4.RDS")

```

When we fit this model we can see the posterior distributions 
of these $\delta$ parameters, which correspond to the 
role the __single previous__ game has in the 
win probability for the next game.

```{r load the stan fit}

fit4 <- readRDS("../model_fits/model4.RDS")

mcmc_hist(fit4$draws(c("delta"))) +
  labs(title = "Effect of Previous Game")

mcmc_hist(fit4$draws(c("beta"))) +
  labs(title = "Effect of Previous n games, excluding most recent game")

```

We can then again compare these estimated effects. Most of these $\delta$
have posterior means between $[-0.2, 0.2]$. 
$\delta=0.2$ corresponds to winning the previous game leading to an 
increase in the win probability of 5%, while holding all 
other variables fixed.

It is maybe not immediate how to compare the effect sizes for $\beta$ and 
$\delta$ here.

```{r check whether positive effects}
#| echo: true

## the beta parameters (previous n, not last game)
prob_positive(fit4)

## the delta parameters (single game) 
prob_positive(fit4, param = "delta")
```

This seems to indicate that the previous game on its own does not have as clear
an effect, but that the streak of previous games shows stronger evidence
for a winner effect.

We will also fit another model, using only
the previous game as history, to this same dataset.

```{r model 5 bullet 60}
#| eval: false

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  prev_game = hist_data_init$prev_game)

stan_file <- "../model5.stan"

mod <- cmdstan_model(stan_file)

fit5 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit5$save_object(file = "../model_fits/model5.RDS")

```

We can see the estimated effects from this model below.

```{r}
fit5 <- readRDS("../model_fits/model5.RDS")

mcmc_hist(fit5$draws(variables = c("delta", "mu3", "tau3", "gamma1", "gamma2")))
prob_positive(fit5$draws(), param = "delta")
```

The estimated effect sizes for $\delta$ are very similar to those estimated
under the previous model, indicating that including the 
longer history provides an improvement over just the previous game 
(particularly given the model comparison below).

## Consider all bullet games

Previously we considered only 60 second bullet games. We wish to 
consider if there is evidence for this effect across all bullet games.

```{r get all bullet games}

all_bullet <- lichess_data %>% 
  filter(Event == "Rated Bullet game")

tidy_games <- map_dfr(top_players, get_hist, all_bullet, prev_n = 10) %>% 
  as_tibble()

tidy_games

stan_data <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
  filter(focal_result != 0.5)

```

```{r refit model 3 for all bullet}
#| eval: false

stan_data_ave <- list(N = nrow(stan_data),
                  J = length(top_players),
                  y = stan_data$focal_result,
                  id = stan_data$focal_id,
                  colour = stan_data$focal_white,
                  elo = stan_data$elo_diff,
                  win_prop = stan_data$ave_prop)

stan_file <- "../model3.stan"

mod <- cmdstan_model(stan_file)

fit3_bullet <- mod$sample(data = stan_data_ave,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit3_bullet$save_object(file = "../model_fits/model3_n10_all_bullet.RDS")
```

```{r}
fit3_bullet <- readRDS("../model_fits/model3_n10_all_bullet.RDS")

mcmc_hist(fit3_bullet$draws(c("beta", "mu1", "tau1", "gamma1", "gamma2")))

prob_positive(fit3_bullet)
```

How do these estimates compare to those with only the 60 second games?
The model with all bullet games has all posterior means greater than 0.
There is not much of an overall change between the estimated mean
effect sizes between the two datasets.
Using all bullet games also leads to 
smaller posterior standard deviations for the
winner effect size for all players.

```{r refit model 4 for all bullet games}
#| eval: false

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

stan_file <- "../model4.stan"

mod <- cmdstan_model(stan_file)

fit4 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit4$summary()

fit4$save_object(file = "../model_fits/model4_n10_all_bullet.RDS")

```

```{r examine this model fit}
fit4 <- readRDS("../model_fits/model4_n10_all_bullet.RDS")

mcmc_hist(fit4$draws(variables = c("beta", "mu1", "tau1", "gamma1", "gamma2")))

mcmc_hist(fit4$draws(variables = c("delta", "mu3", "tau3")))

prob_positive(fit4)
prob_positive(fit4, param = "delta")
```

#### Just previous game

We could also only use the previous game as a predictor, 
without using the rest of the history.

```{r}
#| eval: false

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  prev_game = hist_data_init$prev_game)

stan_file <- "../model5.stan"

mod <- cmdstan_model(stan_file)

fit5 <- mod$sample(data = stan_data_hist,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

fit5$save_object(file = "../model_fits/model5_n10_all_blitz.RDS")

```

```{r}
fit5 <- readRDS("../model_fits/model5_n10_all_blitz.RDS")

mcmc_hist(fit5$draws(variables = c("delta")))

prob_positive(fit5, param = "delta")

```

Fitting using just the previous game as a predictor seems to give a reasonable
model, with a similar interpretation as when we include the history also.

For this model, a posterior mean value for delta of 0.25 corresponds to 7%
increase in the win probability having won the previous game compared 
to having lost it.

Not clear if this difference is due to excluding the win
history or due to using a larger/different dataset.

## Model Checking

We now have a bunch of variations on the original model. We have fit these
on both only 60 second games and all bullet games.
We do some initial model checking, using only the
smaller data initially (to speed up the computation).

As a model check, we use each of the posterior draws to
generate corresponding simulated outcomes, for each
of the games, given the covariates.
We can then use this to get a posterior predictive distribution
for the number of games won by each of the players, which 
we show in the histograms below. Each 
red vertical line corresponds to the true
number of games won by a player.

We see that the true values are very plausible under each
of these models and provide one reasonable sanity check.

```{r get data for this section}
tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>% 
  as_tibble()
## For model 3

init_data <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>% 
  filter(focal_result != 0.5)

stan_data_ave <- list(N = nrow(init_data),
                  J = length(top_players),
                  y = init_data$focal_result,
                  id = init_data$focal_id,
                  colour = init_data$focal_white,
                  elo = init_data$elo_diff,
                  win_prop = init_data$ave_prop)

## For model 4 and model 5

hist_data_init <- tidy_games %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>% 
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         focal_win_prop, elo_diff, focal_result) %>% 
  group_by(focal_id) %>% 
  mutate(ave_prop = lag(focal_win_prop, n = 2, default = 0) - 
           mean(focal_win_prop),
         prev_game = lag(focal_result, default = 0)) %>% 
  filter(focal_result != 0.5)

stan_data_hist <- list(N = nrow(hist_data_init),
                  J = length(top_players),
                  y = hist_data_init$focal_result,
                  id = hist_data_init$focal_id,
                  colour = hist_data_init$focal_white,
                  elo = hist_data_init$elo_diff,
                  win_prop = hist_data_init$ave_prop,
                  prev_game = hist_data_init$prev_game)

```

```{r load single model at a time for model checking}
fit3_ave <- readRDS("../model_fits/init_model_n10.RDS")

fit3_samples <- as_draws_df(fit3_ave$draws())

fit3_samp <- fit3_samples %>% 
  select(!starts_with(c("log_lik", "yrep")))

y_rep <- fit3_samples %>% select(starts_with("y_rep"))

rm(fit3_samples)

### then want to join this with the correct input, i.e player, etc

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_ave$id)

## what is the check here? number of games won by a player for each of
## the posterior samples, compared to the number of games the actually won

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

## then stick the true number of games on top of this and see how
## it looks

orig_data <- tibble(outcome = stan_data_ave$y,
                    focal_id = stan_data_ave$id)

orig_games_won <- orig_data %>% 
  group_by(focal_id) %>% 
  summarise(games_won = sum(outcome))

games_won %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~focal_id, scales = "free") +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  labs(title = "Model 3")

## could get posterior predictive p-values here but they should all 
## be fine, looks pretty decent,
## smallest p-value is 0.06 here for id = 16

```

```{r repeat for model 4}

fit4 <- readRDS("../model_fits/model4.RDS")

fit4_samples <- as_draws_df(fit4$draws())

y_rep <- fit4_samples %>% select(starts_with("y_rep"))

rm(fit4_samples)

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_hist$id)

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

games_won %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~focal_id, scales = "free") +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  labs(title = "Model 4")

## this model looks event better than model 3

```

```{r repeat this posterior check for model 5}

fit5 <- readRDS("../model_fits/model5.RDS")

fit5_samples <- as_draws_df(fit5$draws())

y_rep <- fit5_samples %>% select(starts_with("y_rep"))

rm(fit5_samples)

y_rep_mod <- y_rep %>% 
  rownames_to_column() %>%  
  pivot_longer(-rowname) %>% 
  pivot_wider(names_from=rowname, values_from=value) %>% 
  ## remove the draws and chains here
  mutate(focal_id = stan_data_hist$id)

games_won <- y_rep_mod %>% 
  pivot_longer(cols = `1`:`4000`, names_to = "draw", values_to = "y") %>% 
  group_by(focal_id, draw) %>% 
  summarise(games_won = sum(y))

games_won %>% 
  ggplot(aes(x = games_won)) +
  geom_histogram() +
  facet_wrap(~focal_id, scales = "free") +
  geom_vline(data = orig_games_won, 
             mapping = aes(xintercept = games_won), col = "red") +
  labs(title = "Model 5")

```

Model 5 appears to look as good as model 4, in terms
of posterior predictive distribution for the output.

### Model comparison

Finally, we also consider some model comparison
between these 3 proposed models. Here we
compute the leave one out cross validation.
This allows us to compare across the models, with
the general rule of thumb being a difference 
of more than 2 standard errors indicates a better model.

```{r}
#| cache: true
comp <- loo_compare(fit3_ave$loo(), fit4$loo(), fit5$loo())

comp

## this indicated model 2 slightly better than model 4, clearly better than
## model 3

```

Here we see that the second model, including both the previous game
and
the $n$ games before that, is an improvement over the other 
two models. The model which considers only the single previous game
is closer.

## Further Ideas

- What about some sort of mixture of the previous game and the others before 
that?

- Some sort of decaying effect?

- Is there an interaction between the winner effect and the ability
of the player, in terms of their ELO score?