### 2023-05-23 owen/Write Ups/03_existing_models.qmd
writing some code to get the versions of the models James shared somewhat up and running, have it ready to fit them on the cluster also

---
title: "Model Checking"
author: Owen G. Ward
date: 04/17/2023
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## Model Comparison

We want to compare these potential models, which is very similar
to the model we fit previously.
The only change is that we now need the history for
both players, not just the focal player.

```{r load in the initial data}
#| cache: true

lichess_data <- readRDS("../../rdata/lichess_pilot.RData")

bullet_60 <- lichess_data %>% 
  filter(Event == "Rated Bullet game") %>% 
  filter(TimeControl == "60+0") %>% 
  mutate(game_id = row_number())

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white, game_id) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))s
  hist_games
}

focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18)
## if don't use all focal players here then removing data because they won't be
## involved

top_players <- names(focal_players)

## then get the opponents of these players

full_opponents <- bullet_60 %>% 
  filter(Username %in% top_players) %>% 
  select(White, Black)

all_opps <- c(full_opponents$White, full_opponents$Black)

opponents <- unique(all_opps)

## then need to get the history of these players, but maybe need
## to store game id
## can just match using date and time I guess?
## need to modify the function then to account for this

get_hist_opp <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(opp_white = ifelse(user == White, 1, 0)) %>% 
    select(White:BlackElo, opp_white, game_id) %>% 
    mutate(opp_result = case_when(
      (opp_white == 1 & Result == "1-0") ~ 1,
      (opp_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
      ))

  ngames <- nrow(hist_games)

  if(ngames < prev_n) {
    hist_games <- hist_games %>% 
    mutate(opp_win_prop = opp_result/(1:ngames))
  }
  else{
    hist_games <- hist_games %>% 
      mutate(opp_win_prop = c(cumsum(opp_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(opp_result, n = prev_n)))
  }
  hist_games$player <- user
  hist_games
}
## so now the history is the history corresponding to the name
## in the "player" column

remaining_opponents <- opponents[!(opponents %in% top_players)]

opp_hist <- map_dfr(remaining_opponents, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble()

## so this gives the history of the opponents for each of their 
## games, can use this to then first get their previous result, etc,
## then join this with the other results

prev_opp <- opp_hist %>% 
  group_by(player) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_opp_result = lag(opp_result, default = 0), 
         prev_opp_prop = lag(opp_win_prop, default = 0)) %>%
  select(game_id, player, prev_opp_result, prev_opp_prop)

tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 5) %>% 
  as_tibble()

all <- tidy_games %>% 
  left_join(prev_opp, by = "game_id") %>% 
  mutate(focal = ifelse(focal_white == 1, White, Black))
## this looks more reasonable now
## this has the history of the opponent included, not the history 
## of the focal player yet

final_data <- all %>% 
  group_by(focal) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(prev_focal_result = lag(focal_result, default = 0),
         prev_focal_prop = lag(focal_win_prop, default = 0)) %>% 
  ungroup()

## then tidy this up for stan

```

This object `prev_game_both` should have the previous result of both players in
each game, if that game is in the dataset.
Similarly, `final_prev` has the games played by the focal players, along 
with their previous result and their opponents previous result, if that
game was in the original dataset used (all 60 second bullet games here).

Then convert this into data for stan.

```{r get stan data for simple model}

final_data

stan_prev <- final_data %>% 
  mutate(WhiteElo = as.numeric(WhiteElo), 
         BlackElo = as.numeric(BlackElo)) %>% 
  mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%
  mutate(elo_diff = ifelse(focal_white == 1, 
                           WhiteElo - BlackElo, BlackElo - WhiteElo)) %>% 
  mutate(focal_id = match(focal_user, top_players)) %>% 
  select(focal_user, focal_id, focal_white, 
         prev_focal_result, prev_opp_result, focal_result)

stan_data <- list(N = nrow(stan_prev),
                  J = length(top_players),
                  y = stan_prev$focal_result,
                  id = stan_prev$focal_id,
                  focal_prev = stan_prev$prev_focal_result,
                  opp_prev = stan_prev$prev_opp_result)

```

```{r fit the simple stan model}
#| eval: false

## remember no effect for non focal players in this at all 
## at the moment

stan_file <- "../Comp1.stan"

mod <- cmdstan_model(stan_file)

fit1_simple <- mod$sample(data = stan_data,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)

## this fits at least now...

fit1_simple$summary()
mcmc_hist(fit1_simple$draws(c("alpha", "beta")))

```

## Notes on these models

- How does the choice of focal player make a difference in these models?

```{r thinking about the choice of focal}

table(bullet_60$Username)

tidy_bullet <- bullet_60 %>% 
  select(White:BlackElo, Username, game_id) %>% 
  mutate(opponent = ifelse(Username == White, Black, White),
         focal_white = ifelse(Username == White, 1, 0))

all_players <- c(tidy_bullet$Username, tidy_bullet$opponent)
players <- unique(all_players)

## then just get all the history for this
all_hist <- map_dfr(players, get_hist_opp, bullet_60, prev_n = 5) %>% 
  as_tibble() %>% 
  select(game_id:player)

## this should have each game appear twice, for each of the two players
## involved, so need to figure out how to handle that best
## left join with tidy games matching player to focal and then
## to opponent, rename the columns as needed

tidy_stan <- tidy_bullet %>% 
  left_join(all_hist, by = c("game_id", "Username" = "player")) %>% 
  arrange(UTCDate, UTCTime) %>% 
  rename(focal_result = opp_result, focal_win_prop = opp_win_prop) %>% 
  left_join(all_hist, by = c("game_id", "opponent" = "player")) %>% 
  as_tibble()

## this is all good, just need to lag both and add them in,
## and also get the ids for the stan code

## getting the overall win probabilities for each player also

player_avg <- all_hist %>% 
  group_by(player) %>% 
  summarise(overall_avg = mean(opp_result))

## then just join this in also to use it

```

```{r getting the final data for potential stan model}

final_data <- tidy_stan %>% 
  left_join(player_avg, by = c("Username" = "player")) %>% 
  rename(focal_avg = overall_avg) %>% 
  left_join(player_avg, by = c("opponent" = "player")) %>% 
  rename(opp_avg = overall_avg) %>% 
  group_by(Username) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(focal_prev_result = lag(focal_result, default = 0)) %>% 
  ungroup() %>% 
  group_by(opponent) %>% 
  arrange(UTCDate, UTCTime) %>% 
  mutate(opp_prev_result = lag(opp_result, default = 0)) %>% 
  ungroup() %>% 
  mutate(focal_id = match(Username, players),
         opp_id = match(opponent, players)) %>% 
  filter(focal_result != 0.5)

```

```{r}

final_data

stan_data_all <- list(N = nrow(final_data),
                      J = length(players),
                      y = final_data$focal_result,
                      focal_id = final_data$focal_id,
                      opp_id = final_data$opp_id,
                      focal_prev = final_data$focal_prev_result,
                      opp_prev = final_data$opp_prev_result,
                      focal_avg = final_data$focal_avg,
                      opp_avg = final_data$opp_avg)

saveRDS(stan_data_all, file = "../cluster_scripts/stan_data.RDS")

stan_file <- "../Comp2.stan"

mod2 <- cmdstan_model(stan_file)

fit2 <- mod2$sample(data = stan_data_all,
                    seed = 123,
                    chains = 4,
                    parallel_chains = 4,
                    refresh = 100,
                    iter_sampling = 500,
                    iter_warmup = 100)

```