### 2023-12-08 owen/Write Ups/04_model_checking.qmd
update, add this file to github

---
title: "Model Checking"
author: Owen G. Ward
date: today
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
bibliography: refs.bib
---

The aim of this document is to summarise the current version 
of the model and describe some of the ways we have assessed the
suitability of this model to our data. The
aim is for this document to be self contained.

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 8)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(here)

options(mc.cores = parallel::detectCores())

theme_set(theme_bw())

prob_positive <- function(stan_draws, param = "beta") {
  all_draws <- as_draws_df(stan_draws)
  ndraws <- nrow(all_draws)
  as_tibble(all_draws) %>% 
  select(starts_with(param)) %>% 
  apply(2, function(x) sum(x>0)/ndraws)
}

```

## The current model

Based on @ding2021nhl, an initial model proposed was of the form

$$
P(y_{ij} = 1) = \frac{1}{1 + \exp(-(\alpha_j + \beta_j x_{ij} + \gamma z_{ij}))}
$$

where this is the probability of the **j-th focal player** 
winning their **i-th game**. Here:

- $\alpha_j$ is a player level random effect
- $\beta_j$ is a player level random effect, accounting for the win ratio $x_{ij}$
of the focal player over their previous $n$ games
- $\gamma$ is a fixed effect of game level coefficients (colour, difference
in ELO ranking)

We then partially pool the $\alpha$ and $\beta$ coefficients, so

$$
\alpha_j \sim \mathcal{N}(\mu_2, 1)
$$
$$
\mu_2 \sim \mathcal{N}(0, 1)
$$
$$
\tau_2 \sim \mathcal{N}(0, 1)
$$

and similarly

$$
\beta_j \sim \mathcal{N}(\mu_1, 1)
$$
$$
\mu_1 \sim \mathcal{N}(0, 1)
$$
$$
\tau_1 \sim \mathcal{N}(0, 1)
$$

We rescale the win ratio $x_{ij}$, by the average win ratio of
player $j$ across all 
games they play in the dataset.
This means $\beta_j$ captures how a players win
probability changes as their historic performance deviates from 
their overall average win ratio.

For $z_{ij}$ we incorporate 2 covariates, namely the colour of 
the focal player and the ELO difference between the focal 
player and their opponent. So that means we have 
$\gamma_1$, which tells us how the win probability changes in going
from black to white,
and $\gamma_2$ gives how the win probability changes as the focal player
has a better ELO than their opponent.

__These predictors are not currently rescaled__.

## Fitting this to Lichess Data

We fit this to some of the large lichess data. We consider only a subset 
of the players in a given rating range, due to time constraints.

### 1700-1900 Players

We fit this to games from 25 players in the 1700-1900 range. 
We looked only at standard rated bullet games (no tournaments, etc),
which was 445k games, ranging from 500 - 50k games per player.
Fitting this model take over 12 hours on a computing cluster.

This results in the following estimated parameters.

```{r}
fit <- readRDS(here("results/", "lichess1700-1900/", 
                    "all_rated_bullet_model3.RDS"))

mcmc_hist(fit$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

```

- Here the mean for $\mu_2$ is -0.0651, corresponding to a 48% chance of 
winning as black against an equally ranked opponent. This parameter
is not clearly different from 0 here. Posterior SD is large.

- $\gamma_1$ is estimated to be about 0.115, indicating that playing 
as white increases the win probability by 3%, holding other parameters 
fixed. Posterior SD is very small.

- $\gamma_2$ indicates that a 10 point ELO difference increases the win 
probability by about 1%, while a 100 ELO score difference 
increases the win probability by about 9%. Posterior SD is very small.

- The global estimate for $\mu_1$, the global winner effect, is 
around 0. Posterior SD is large here.

```{r get names 1700-1900}
#| cache: true

## need to get user names here, from the data again

data_path <- here("box_data/lichess1700-1900/")

files <- list.files(data_path)
files <- files[1:25]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
    )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r winner-effects-17-19}
#| fig.height = 7

players <- users
names(players) <- paste0("beta[", 1:length(users), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")
```

Here we see that there is a range of plausible winner effects for 
different players.

- Some have posterior distributions which are only positive, some 
are only negative and some are centered around or near 0.

-The largest posterior mean here is 1.32. This is saying that if on
average you win 50% of your games, but have instead won 10 in a row, your
probability of winning the next would increase by ~16% over that baseline.
However most estimated effects are small, and most of these posterior
distributions contain 0.

```{r indiv-effects-17-19}
#| fig.height = 6

names(players) <- paste0("alpha[", 1:length(users), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

- Most of the individual player effects are somewhat negative but
close to 0.

- Many are quite similar to the global parameter here.

### 2000-2200 Players

We fit this model also to 10 players rated between 2000 and 2200. 
With the same restrictions on the games this gives us 285k games.
We have between 1200 and 89k games for each of these 10 players.

```{r}

fit2 <- readRDS(here("results/", "lichess2000-2200/", 
                    "all_rated_bullet_model.RDS"))

mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")

```

```{r get-player-names}
#| cache: true

data_path <- here("box_data/lichess2000-2200/")

files <- list.files(data_path)
files <- files[1:10]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

get_hist <- function(user, games, prev_n) {
  hist_games <- games %>% 
    filter(White == user | Black == user) %>% 
    arrange(UTCDate, UTCTime) %>% 
    mutate(focal_white = ifelse(Username == White, 1, 0)) %>% 
    select(White:BlackElo, focal_white) %>% 
    mutate(focal_result = case_when(
      (focal_white == 1 & Result == "1-0") ~ 1,
      (focal_white == 0 & Result == "0-1") ~ 1,
      (Result == "1/2-1/2") ~ 0.5,
      .default = 0
    )) %>% 
    mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)), 
                              roll_mean(focal_result, n = prev_n)))

  hist_games
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users_20_22 <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r}

players <- users_20_22
names(players) <- paste0("beta[", 1:length(users_20_22), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit2$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")

```

```{r}
names(players) <- paste0("alpha[", 1:length(users_20_22), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit2$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

- Here the effect of a difference in ranking appears to be slightly larger.
This leads to a bigger change in the win probability when the difference
in ranking between the two players is larger.

- The posterior standard deviations of the $\mu$ parameters is slightly
larger in this model, but this could be explained by the smaller number of 
games.

- A smaller sample, but half of these posterior distributions 
of the winner effects contain 0.

### 2300-2500 Players

We look at 8 players rated between 2300 and 2500. This gives us 160k games,
with players playing 4k-58k games.

```{r}
fit3 <- readRDS(here("results/", "lichess2300-2500/", 
                    "all_rated_bullet_model.RDS"))

mcmc_hist(fit2$draws(variables = c("mu1", "mu2", "gamma1", "gamma2"))) +
  labs(title = "Global Parameters",
       subtitle = "mu1 is Winner Effect, mu2 is Player Effect, Gamma1 is colour,
       Gamma2 is ELO Difference")
```

```{r get users 2300-2500}
#| cache: true

data_path <- here("box_data/lichess2300-2500/")

files <- list.files(data_path)
files <- files[1:8]

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

## restrict to rated rapid and shorter here
## this also removes the NAs, which makes sense

small_data <- lichess_data %>% 
  # filter(Event == "Rated Bullet game") %>% 
  # filter(TimeControl == "60+0") %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event)) #%>% 
  # filter(!grepl("Classical|Correspondence", Event))

users_23_25 <- unique(small_data$Username)

rm(lichess_data)
rm(small_data)

```

```{r}

players <- users_23_25
names(players) <- paste0("beta[", 1:length(users_23_25), "]")

player_labels <- as_labeller(players)

mcmc_hist(fit3$draws("beta"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Player Winner Effects")

```

```{r}
names(players) <- paste0("alpha[", 1:length(users_23_25), "]")
player_labels <- as_labeller(players)
mcmc_hist(fit3$draws("alpha"),
          facet_args = list(labeller = player_labels)) +
  labs(title = "Individual Player Effects")
```

## Other Questions

Do focal players play each other in this data?

```{r}
data_path <- here("box_data/lichess1700-1900/")

files <- list.files(data_path)
files <- files[1:25]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

users <- unique(lichess_data$Username)

overlap_data <- lichess_data %>% 
  filter(White %in% users & Black %in% users) %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event))

rm(lichess_data)

overlap_data %>% group_by(Username) %>% 
  tally() %>% 
  arrange(-n)
```

Certain players have a large amount of overlap, looking only at the games we
consider in the analysis. How should these games be accounted for?

```{r check for 2000-2200 games also}
#| eval: false

## happens here also, but not as often for this sample

data_path <- here("box_data/lichess2000-2200/")

files <- list.files(data_path)
files <- files[1:10]

read_player <- function(path, file){
  dat <- read_csv(file = paste0(data_path, file),
                  col_types = cols(UTCDate = col_date("%Y.%m.%d"),
                                   WhiteTitle = col_character(),
                                   BlackTitle = col_character(),
                                   WhiteElo = col_character(),
                                   BlackElo = col_character(),
                                   FEN = col_character())) %>% 
    select(Username, Event, White, Black, Result, UTCDate, UTCTime, 
           WhiteElo, BlackElo, Variant, TimeControl, Termination) %>% 
    mutate(WhiteElo = parse_number(if_else(WhiteElo == "?", NA, WhiteElo)),
           BlackElo = parse_number(if_else(BlackElo == "?", NA, BlackElo)))
  dat
}

lichess_data <- files %>% 
  map_dfr(~read_player(data_path, .x))

users <- unique(lichess_data$Username)

overlap_data <- lichess_data %>% 
  filter(White %in% users & Black %in% users) %>% 
  filter(Variant == "Standard") %>% 
  filter(grepl("Rated Bullet game", Event))

rm(lichess_data)

overlap_data %>% group_by(Username) %>% 
  tally() %>% 
  arrange(-n)

```

### Other Comments

- Model checking is hard for these large datasets

- Not easy to do the permutation fits for all games

- Hard to simulate fake data also

- Are we ignoring any covariates which could influence the model here?

- How do these effects change if we only look at games which were played 
directly before?

<!-- Below some ideas about posterior predictive checking, hard to currently  -->
<!-- consider for the larger datasets we're working with. -->

<!-- ## Posterior Predictive Checking -->

<!-- ## Permuting the Original Data -->

<!-- We can also use permutations of the data to examine how -->
<!-- robust these results are. -->

<!-- ### Global Permutation -->

<!-- The simplest permutation of our model would be to simply permute -->
<!-- all results at random and then process the data and refit this model. -->

<!-- ```{r} -->
<!-- perm_path <- here("owen", "cluster_scripts", "perm_outputs", "perm_outputs") -->

<!-- sims <- list.files(path = perm_path) -->

<!-- perm_results <- sims %>%  -->
<!--   map_dfr(~readRDS(here("owen", "cluster_scripts", -->
<!--                         "perm_outputs", "perm_outputs", -->
<!--                         .x))) %>%  -->
<!--   select(variable, mean, median, sim_id) -->

<!-- ## global parameters here -->

<!-- fit3_ave <- readRDS("../model_fits/init_model_n10.RDS") -->

<!-- orig <- fit3_ave$summary(c("alpha", "beta", "gamma1", "gamma2", "mu1", "mu2", -->
<!--                    "tau1", "tau2"))  -->

<!-- rm(fit3_ave)  -->

<!-- perm_results %>%  -->
<!--   filter(variable == "mu1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of Mu_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "mu1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "mu1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "mu2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of Mu_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "mu2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "mu2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- ## try it also for some of the winner effects -->

<!-- perm_results %>%  -->
<!--   filter(variable == "gamma1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of gamma_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "gamma_1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "gamma1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "gamma2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of gamma_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "gamma_2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "gamma2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "tau1") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of tau_1", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "tau1") + -->
<!--   geom_vline(data = orig %>%  -->
<!--                filter(variable == "tau1"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- perm_results %>%  -->
<!--   filter(variable == "tau2") %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = "Permutations of tau_2", -->
<!--        subtitle = "True data appears different", -->
<!--        x = "tau_2") + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == "tau2"), -->
<!--              aes(xintercept = mean), col = "red") -->

<!-- ``` -->

<!-- This quick permutation seem to indicate that there is evidence for these  -->
<!-- effects, which is not present when the results are randomly permuted  -->
<!-- instead. -->

<!-- Now we can look at the permutation distributions of $\alpha$. These are -->
<!-- centered at $\mu_2$ with standard deviation $\tau_2$. -->
<!-- As expected, all permutation distributions are essentially  -->
<!-- draws from the the global mean player effect, which corresponds to some  -->
<!-- overall number of games won. -->
<!-- Most of the estimates fall in these permutation distributions, essentially -->
<!-- indicating that the individual effect is the same as the global effect for these -->
<!-- players.  -->
<!-- For the players with a smaller effect, they are less likely to win -->
<!-- then the overall average, while players with a larger effect are more likely to -->
<!-- win. -->

<!-- I suspect the jumpy nature of these permutation distributions would -->
<!-- disappear if more permutations were considered. -->

<!-- ```{r repeat for all alpha} -->

<!-- alpha_var <- paste0("alpha[", 1:18, "]") -->

<!-- # alpha_var -->

<!-- for(i in seq_along(alpha_var)){ -->
<!--   plot_curr <- perm_results %>%  -->
<!--   filter(variable == alpha_var[i]) %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = paste0("Permutations of ", alpha_var[i]), -->
<!--        subtitle = "True data appears different", -->
<!--        x = alpha_var[i]) + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == alpha_var[i]), -->
<!--              aes(xintercept = mean), col = "red") -->
<!--   print(plot_curr) -->
<!-- } -->

<!-- ``` -->

<!-- Given this global permutation of all game results, if we look at the  -->
<!-- winner effect, we should expect to not see any evidence for these under -->
<!-- the permutation distribution. The individual effects have mean $\mu_1$  -->
<!-- and sd $\tau_1$. Note that these $\beta$ estimates under the  -->
<!-- original data, which correspond to to winner and loser effects, appear -->
<!-- to have stronger evidence under this permutation. Almost all of these are  -->
<!-- positive and the one negative estimate is likely not significant, based -->
<!-- on the posterior distribution anyway. -->

<!-- ```{r repeat for all beta} -->

<!-- beta_var <- paste0("beta[", 1:18, "]") -->

<!-- # alpha_var -->

<!-- for(i in seq_along(beta_var)){ -->
<!--   plot_curr <- perm_results %>%  -->
<!--   filter(variable == beta_var[i]) %>%  -->
<!--   ggplot(aes(mean)) + -->
<!--   geom_histogram() + -->
<!--   labs(title = paste0("Permutations of ", beta_var[i]), -->
<!--        subtitle = "True data appears different", -->
<!--        x = beta_var[i]) + -->
<!--   geom_vline(data = orig %>%  -->
<!--              filter(variable == beta_var[i]), -->
<!--              aes(xintercept = mean), col = "red") -->
<!--   print(plot_curr) -->
<!-- } -->

<!-- ``` -->

<!-- Are the estimates that are not centered around 0 explained by the  -->
<!-- overall proportion of wins in the dataset? When you permute all the results -->
<!-- you essentially get that 48% of games are won by a focal player playing -->
<!-- as black. This is basically $1/(1+\exp(-1(-0.1)))$. Note that the  -->
<!-- proportion is higher in the original data. -->

<!-- ```{r} -->
<!-- #| eval: true -->
<!-- #| echo: false -->

<!-- lichess_data <- readRDS("../../rdata/lichess_pilot.RData") -->

<!-- bullet_60 <- lichess_data %>%  -->
<!--   filter(Event == "Rated Bullet game") %>%  -->
<!--   filter(TimeControl == "60+0") %>%  -->
<!--   mutate(game_id = row_number())  -->

<!-- get_hist <- function(user, games, prev_n) { -->
<!--   hist_games <- games %>%  -->
<!--     filter(White == user | Black == user) %>%  -->
<!--     arrange(UTCDate, UTCTime) %>%  -->
<!--     mutate(focal_white = ifelse(Username == White, 1, 0)) %>%  -->
<!--     select(White:BlackElo, focal_white) %>%  -->
<!--     mutate(focal_result = case_when( -->
<!--       (focal_white == 1 & Result == "1-0") ~ 1, -->
<!--       (focal_white == 0 & Result == "0-1") ~ 1, -->
<!--       (Result == "1/2-1/2") ~ 0.5, -->
<!--       .default = 0 -->
<!--       )) %>%  -->
<!--     mutate(focal_win_prop = c(cumsum(focal_result[1:(prev_n - 1)])/(1:(prev_n -1)),  -->
<!--                               roll_mean(focal_result, n = prev_n))) -->

<!--   hist_games -->
<!-- } -->

<!-- focal_players <- head(sort(table(c(bullet_60$Username)), decreasing = TRUE), 18) -->

<!-- top_players <- names(focal_players) -->

<!-- tidy_games <- map_dfr(top_players, get_hist, bullet_60, prev_n = 10) %>%  -->
<!--   as_tibble() -->

<!-- init_data <- tidy_games %>%  -->
<!--   mutate(WhiteElo = as.numeric(WhiteElo),  -->
<!--          BlackElo = as.numeric(BlackElo)) %>%  -->
<!--   mutate(focal_user = ifelse(focal_white == 1, White, Black)) %>%  -->
<!--   mutate(elo_diff = ifelse(focal_white == 1,  -->
<!--                            WhiteElo - BlackElo, BlackElo - WhiteElo)) %>%  -->
<!--   mutate(focal_id = match(focal_user, top_players)) %>%  -->
<!--   select(focal_user, focal_id, focal_white,  -->
<!--          focal_win_prop, elo_diff, focal_result) %>%  -->
<!--   group_by(focal_id) %>%  -->
<!--   mutate(ave_prop = lag(focal_win_prop, default = 0) - mean(focal_win_prop)) %>%  -->
<!--   filter(focal_result != 0.5) -->

<!-- init_data %>% -->
<!--   group_by(focal_white, focal_result) %>% -->
<!--   tally() %>% -->
<!--   mutate(prop = n /sum(n)) -->
<!-- #  -->

<!-- ``` -->

<!-- ### Focal Player Permutation -->

<!-- Alternatively, we could restrict our permutations to  -->
<!-- instead be at the level of each focal player. This -->
<!-- means that the total number of games won by each focal -->
<!-- player will be the same, but any temporal -->
<!-- structure to these results will be removed. -->

<!-- ## Questions -->

<!-- - Are we ignoring any covariates which could influence the model here? -->

<!-- - How do these effects change if we only look at games which were played  -->
<!-- directly before? -->