### 2025-01-24 Paper_WriteUp/Paper_Plots.qmd
add some supplement plots

---
title: "Paper Analysis"
author: Owen G. Ward
date: today
format: 
  # html:
  #   toc: true
  #   toc-location: left
  pdf
---

This document will recreate all numeric and graphical
summaries contained in the submitted manuscript and the 
supplementary materials.

```{r setup}
#| include: false

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 7, dev = "png", fig.path = "output/",
                      fig.height = 5, dpi = 600)

library(tidyverse)
library(RcppRoll)
library(cmdstanr)
library(posterior)
library(bayesplot)
library(loo)
library(scales)
library(ggridges)
library(here)
library(ggdist)
library(viridis)
library(PlayerRatings)
library(ggrepel)
library(ggpubr)
options(mc.cores = parallel::detectCores())
theme_set(theme_bw())

## to load some plot defaults
source(here("analysis", "helper.R"))
source(here("analysis", "plot_templates.R"))

all_data_path <- rep(NA, 4)
all_data_path[1] <- here("box_data/lichess1700-1900/")
all_data_path[2] <- here("box_data/lichess2000-2200/")
all_data_path[3] <- here("box_data/lichess2300-2500/")
all_data_path[4] <- here("box_data/lichessGrandmasters/")
res_data_path <- rep(NA, 4)
res_data_path[1] <- here("results_2025/lichess1700-1900/")
res_data_path[2] <- here("results_2025/lichess2000-2200/")
res_data_path[3] <- here("results_2025/lichess2300-2500/")
res_data_path[4] <- here("results_2025/lichessGrandmasters/")

```

# Section 2: Data

We first get some summary statistics about the overall data. 
We construct these for the 1700-1900 players and GMs.
We then examine the number of consecutive games played 
across both cohorts.

```{r summary statistics 1700-1900}
#| eval: false
## this can be run if interested
path_id <- 1  ## for 1700-1900 
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

## overall summary of all event types
lichess_data |> 
  group_by(Event) |> 
  count() |> 
  arrange(-n)

## what proportion of bullet games are of what time format
lichess_data |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated bullet game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

## repeat for blitz games instead
lichess_data |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated blitz game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

bullet_data <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))

bullet_data |> nrow()
## average number, we remove the two accounts with 1 and 3 games only
## throughout we do not consider those who play less than 10 games
## at a given format when fitting any model
bullet_data |> 
  group_by(Username) |> 
  count() |> 
  filter(n > 10) |> # remove 2 players with <10 games
  ungroup() |> 
  summarise(avg = mean(n))

## for each player, count how many times they start
## a game with 5 minutes of the start of a previous game
## and get the proportion across all focal players
 bullet_data |>  
   group_by(Username) |> 
   dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_5, na.rm = TRUE)/ n())

blitz_data <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "180+0") |>
  filter(Variant == "Standard") |>
  filter(grepl("rated blitz game", Event))

## count how many blitz games they play as comparison

blitz_data |> 
  group_by(Username) |> 
  count() |> 
  filter(n > 10) |> # remove players with < 10 games
  ungroup() |> 
  summarise(avg = mean(n))

## repeat for blitz data using 7 minutes as the maximum gap
blitz_data |>  
   group_by(Username) |> 
   dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_7 = ifelse(gap_seconds <= 420, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_7, na.rm = TRUE)/ n())

## examine how many games are played in a session for the bullet data
## look at the streaks, number of games played in a session
outcomes <- bullet_data |> 
  # group_by(Username) |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))

# Apply function to each group
result <- outcomes |> 
  summarise(session_length = list(get_true_sequences(less_3)))

result |> 
  unnest_longer(session_length) |> 
  summarise(med_session = median(session_length), 
            avg = mean(session_length))

## plot the histogram of games per session
result |> 
  unnest_longer(session_length) |> 
  ggplot(aes(session_length)) +
  geom_histogram()

## proportion of sessions with more than 1 game
result |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

## with more than 10 games
result |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

## repeat for blitz games
outcomes_blitz <- blitz_data |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7 = if_else(gap_seconds <= 420, TRUE, FALSE, missing = FALSE))

result_blitz <- outcomes_blitz |> 
  summarise(session_length = list(get_true_sequences(less_7)))

result_blitz |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())
```

```{r summary statistics gm}
#| eval: false
## can be run if interested
## then repeat this analysis for the gms
path_id <- 4  ## for GMs
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data_gm <- files |> 
  map_dfr(~read_player(data_path, .x))

## get proportion of bullet games that are of each time format
lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated bullet game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

## repeat for blitz
lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated blitz game", Event)) |> 
  group_by(TimeControl) |> 
  count() |> 
  arrange(-n) |> 
  ungroup() |> 
  mutate(prop = n/sum(n))

bullet_data_gm <- lichess_data_gm |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))

bullet_data_gm |> nrow()
bullet_data_gm |> 
  group_by(Termination) |>
  tally() |> 
  mutate(prop = n/sum(n))

## average across all focal players
bullet_data_gm |> 
  group_by(Username) |> 
  count() |> 
  ungroup() |> 
  summarise(mean(n))

blitz_data_gm <- lichess_data_gm |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "180+0") |>
  filter(Variant == "Standard") |>
  filter(grepl("rated blitz game", Event))

blitz_data_gm |> 
  group_by(Username) |> 
  count() |> 
  ungroup() |> 
  summarise(mean(n))

## get the average rating of the GMs at the time data
## was collected 
lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated blitz game", Event)) |> 
  group_by(Username) |> 
  arrange(UTCDate, UTCTime) |> 
  slice_tail(n = 1) |> 
  mutate(user_elo = ifelse(Username == White, WhiteElo, BlackElo)) |> 
  dplyr::select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
  ungroup() |> 
  summarise(mean(user_elo))

## look at the proportion of games started within 5 minutes 
## of the previous game
bullet_data_gm |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
  ungroup() |>   ## to get the overall proportion
  summarise(sum(less_5, na.rm = TRUE)/ n())

## look at the streaks, number of games played in a session
outcomes_gm <- bullet_data_gm |> 
  # group_by(Username) |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = if_else(gap_seconds <= 300, TRUE, FALSE, missing = FALSE))

## extract the sessions containing more than 1 game
get_true_sequences <- function(x) {
  rle_x <- rle(x)
  true_lengths <- rle_x$lengths[rle_x$values == TRUE]
  return(true_lengths)
}

# Apply function to each group
result_gm <- outcomes_gm |> 
  summarise(session_length = list(get_true_sequences(less_5)))

result_gm |> 
  unnest_longer(session_length) |> 
  summarise(med_session = median(session_length), 
            avg = mean(session_length))

## plot the histogram of games per session
result_gm |> 
  unnest_longer(session_length) |> 
  ggplot(aes(session_length)) +
  geom_histogram()

## proportion of sessions with more than 1 game
result_gm |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

result_gm |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 10, TRUE, FALSE)) |>  
  summarise(sum(multiple)/n())

## repeat for blitz
outcomes_blitz <- blitz_data_gm |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7_min = if_else(gap_seconds <= 420, TRUE, FALSE, FALSE))

# Apply function to each group
result_blitz <- outcomes_blitz |>
  summarise(session_length = list(get_true_sequences(less_7_min)))

result_blitz |> 
  unnest_longer(session_length) |> 
  ggplot(aes(session_length)) +
  geom_histogram()

result_blitz |> 
  unnest_longer(session_length) |> 
  mutate(multiple = if_else(session_length > 1, TRUE, FALSE)) |> 
  # group_by(Username) |> 
  summarise(sum(multiple)/n())

```

```{r figure 1}
## Sydney's code for this figure

player_cohort_1 <- "sandro2502"

path_id <- 1  ## for 1700-1900 
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

bullet_data_fig_1 <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))
rm(lichess_data)

p1 <- bullet_data |> 
  filter(Username == player_cohort_1)

p1$newELO <- ifelse(p1$White == paste0(player_cohort_1),
                    p1$WhiteElo, p1$BlackElo)
# p1$datetime <- as.POSIXct(paste(p1$UTCDate, p1$UTCTime),
#                           format = "%Y.%m.%d %H:%M:%S")
ltypes <- c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1, 2, 3)
thetitle <- paste0("Player From 1700-1900 Cohort")
linewd <- 1
ylim1 <- 1200
ylim2 <- 3500
ndays <- 1

p1_sorted <- p1 |> arrange(UTCDate, UTCTime)
p1_sorted$event <- 1:nrow(p1_sorted)

tidy_plot1 <- ggplot(p1_sorted, aes(x = event, y = newELO)) +
  # scale_colour_manual(values = getPalette(colourCount)) +
  scale_linetype_manual(values = ltypes) +
  xlab("Games Played") +
  ylab("Lichess Rating") +
  ggtitle(thetitle) +
  ylim(ylim1, ylim2) +
  geom_line(lwd = linewd) +
  theme(
    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
    panel.background = element_blank(),
    plot.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = rel(1.1)),
    text = element_text(color = "gray20", size = 10),
    axis.text = element_text(size = rel(1.0)),
    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
    axis.title.x = element_text(size = rel(1.0), vjust = 0),
    axis.title.y = element_text(size = rel(1.0), vjust = 1),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

player_cohort_gm <- "alireza2003"

## load in the GM data for this
path_id <- 4  ## for GM 
data_path <- all_data_path[path_id]
files <- list.files(data_path)
lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

bullet_data_fig_1_gm <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))
rm(lichess_data)

p2 <- bullet_data_fig_1_gm |> 
  filter(Username == player_cohort_gm)

p2$newELO <- ifelse(p2$White == paste0(player_cohort_gm),
                    p2$WhiteElo, p2$BlackElo)
thetitle <- paste0("Player From GM Cohort")
p2_sorted <- p2 |> arrange(UTCDate, UTCTime)
p2_sorted$event <- 1:nrow(p2_sorted)

tidy_plot2 <- ggplot(p2_sorted, aes(x = event, y = newELO)) +
  # scale_colour_manual(values = getPalette(colourCount)) +
  scale_linetype_manual(values = ltypes) +
  xlab("Games Played") +
  ylab("Lichess Rating") +
  ggtitle(thetitle) +
  ylim(ylim1, ylim2) +
  geom_line(lwd = linewd) +
  theme(
    plot.title = element_text(hjust = 0, vjust = 1, size = rel(1.7)),
    panel.background = element_blank(),
    plot.background = element_blank(),
    panel.grid.major.y = element_line(color = "gray75", linetype = "dotted"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = rel(1.1)),
    text = element_text(color = "gray20", size = 10),
    axis.text = element_text(size = rel(1.0)),
    axis.text.x = element_text(color = "gray20", size = rel(1.0)),
    axis.text.y = element_text(color = "gray20", size = rel(1.0)),
    axis.title.x = element_text(size = rel(1.0), vjust = 0),
    axis.title.y = element_text(size = rel(1.0), vjust = 1),
    # axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none"
  )

tdy1 <- tidy_plot1 + theme_single_y()
tdy2 <- tidy_plot2 + theme_single_y()

tidy_plot <- ggarrange(tdy1, tdy2)

ggsave(paste0(newfilename, ".png"), tidy_plot,
  dpi = 600, height = 5, width = 12
)
ggsave(plot = tidy_plot, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "elo_plot_2_players.png")), dpi = 600,
       height = 5, width = 12)
```

```{r figure 2}
## Adam's code for this figure

gm_players_elo <- list()
gm_session_lengths <- list()
gm_time_diffs <- list()
gm_time_diffs_10 <- list()

gm_players <- unique(bullet_data_gm$Username)

for (i in 1:length(gm_players)) {
  tidy_games_gms <- bullet_data_gm |>
    filter(Username == gm_players[i]) |> # select only one player
    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
    arrange(UTCDateTime) |> # order date-time
    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games

  ## A check to remove duplicated games, if they occur
  # cat("Dupes:", sum(duplicated(tidy_games_gms[1:9])), "\n", sep = " ")
  # # some players have all their games double recorded for some reason
  # tidy_games_gms <- tidy_games_gms[!duplicated(tidy_games_gms[1:9]), ] # remove duplicate rows

  # get players average elo across all games
  gm_players_elo[[gm_players[i]]] <- mean(c(
    tidy_games_gms$WhiteElo[gm_players[i] == tidy_games_gms$White],
    tidy_games_gms$BlackElo[gm_players[i] == tidy_games_gms$Black]
  ))

  # store time diffs between games
  gm_time_diffs[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
    lag(tidy_games_gms$UTCDateTime),
    units = "secs"
  ) # time between current game and previous
  gm_time_diffs_10[[gm_players[i]]] <- difftime(tidy_games_gms$UTCDateTime,
    lag(tidy_games_gms$UTCDateTime, 10),
    units = "secs"
  ) # time between current game and 10th game ago
}
# store in df for plotting
gm_time_diff_df <- data.frame("time_diff" = unlist(gm_time_diffs))
gm_time_diff_df_10 <- data.frame("time_diff" = unlist(gm_time_diffs_10))

## Same distributions as amateurs, so again proceed by assuming all 
## games played within 300 seconds (5 min) are in the same session
for (i in 1:length(gm_players)) {
  tidy_games_gms <- bullet_data_gm |>
    filter(Username == gm_players[i]) |> # select only one player
    mutate(UTCDateTime = ymd_hms(paste(UTCDate, UTCTime))) |> # convert date-time to lubridate type
    arrange(UTCDateTime) |> # order date-time
    mutate(time_diff = UTCDateTime - lag(UTCDateTime)) # time difference between games

  length_of_sessions <- which(tidy_games_gms$time_diff > 300) - 
    lag(which(tidy_games_gms$time_diff > 300), default = 0)
  # game id of first game of session minus game id of first game of previous session

  gm_session_lengths[[gm_players[i]]] <- length_of_sessions # store
}
# store in df for plotting
gm_session_lengths_df <- data.frame(session_length = unlist(gm_session_lengths))

log_time_diff_plot_gm <- ggplot(gm_time_diff_df, aes(x = as.numeric(time_diff))) +
  geom_histogram() +
  scale_x_continuous(
    trans = scales::log_trans(),
    breaks = c(1, 60, 300, 3600, 86400, 31536000),
    labels = c("1 sec", "1 min", "5 min", "1 hour", "1 day", "1 year")
  ) +
  xlab("Time Between Games (Log Scale)") +
  geom_vline(xintercept = 300, colour = "red") +
  ylab("Frequency") #+ ggtitle("Time Difference Between Current Game and Previous Game for Amateurs")
fig2a <- log_time_diff_plot_gm + theme_single()

ggsave(plot = fig2a, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "time_between_games_gm.png")), dpi = 600,
       height = 5, width = 7)

session_length_plot_gm <- ggplot(gm_session_lengths_df, aes(x = session_length)) +
  geom_histogram() +
  scale_x_continuous(
    trans = pseudo_log_trans(sigma = 2.36),
    breaks = c(1, 5, 10, 100),
    labels = c("1 game", "5 games", "10 games", "100 games"),
    limits = c(0, 389)
  ) +
  xlab("Number of Games in a Session (Log Scale)") +
  ylab("Frequency") #+ ggtitle("Number of Games per Session for GMs")
fig2b <- session_length_plot_gm + theme_single()

ggsave(plot = fig2b, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "session_length_gm.png")), dpi = 600,
       height = 5, width = 7)
```

# Section 4: The Fitted Model

## 1700-1900 Previous Game

### Global Parameters

```{r figure 3 left - plot global parameters for 17-19}
data_path <- here("results_2025/lichess1700-1900")

fit_17_19_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

facet_labels <- as_labeller(c(
  mu_beta = "mu[beta]",
  gamma1 = "gamma[1]",
  gamma2 = "gamma[2]",
  sigma_1 = "sigma[1]",
  `tau[1]` = "tau[1]",
  `tau[2]` = "tau[2]",
  sigma_g1 = "sigma[g[1]]",
  sigma_g2 = "sigma[g[2]]"
), label_parsed)

global_17_19_bullet_prev <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
  facet_wrap(~name, ncol = 4, scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3))

global_17_19_bullet_prev
ggsave(plot = global_17_19_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_global.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients 17-19}
#| eval: false
gamma_ests <- fit_17_19_bullet_prev$summary(variables = 
                                              c("gamma1",
                                                "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests[1])))

1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests[2])))
```

### Winner and Player Effects

```{r player level parameters for 17-19}
#| eval: false

player_id <- as_tibble(bullet_17_19) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_17_19_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

## see what the smallest and largest posterior mean is 
## for the winner effects
player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_max(order_by = post_mean, n = 5)

player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_min(order_by = post_mean, n = 5)

## see how many contain 0
player_pars |> 
  group_by(player_name, param) |> 
  filter(param == 2) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  filter(contain_zero == 0)
```

```{r figure 4 left}
winner_plot_17_19 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "1700-1900 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = winner_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "17_19_prev_winner.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 left - plot player effects}
#| eval: false
indiv_plot_17_19 <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank()) +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = indiv_plot_17_19, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S4_left.png")), dpi = 600,
       height = 7, width = 7)
```

## GM previous game

### Global Parameters

```{r figure 3 right - global parameters for gm data}
data_path <- here("results_2025/lichessGrandmasters")

fit_gm_bullet_prev <- 
  readRDS(file = paste0(data_path,"/all_rated_bullet_model_prev_n1.RDS"))

bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

global_gm_bullet_prev <- fit_gm_bullet_prev$draws(c("mu_beta",
                                                          "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"sigma_g2") |> 
  ggplot(aes(value)) +
  geom_histogram() +
   facet_wrap(~name, ncol = 4, scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank()) +
  theme_4col() +
  scale_x_continuous(breaks = breaks_pretty(n = 2))

global_gm_bullet_prev
ggsave(plot = global_gm_bullet_prev, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_global.png")), dpi = 600,
       height = 5, width = 7)
```

```{r interpret the estimated coefficients gm}
#| eval: false
gamma_ests_gm <- fit_gm_bullet_prev$summary(variables = c("gamma1",
                                                          "gamma2")) |> 
  pull(mean)
## compute change in win probability for gamma1
1/(1 + exp(-1 * (0)))
1/(1 + exp(-1 * (gamma_ests_gm[1])))
## change in win probability for elo difference
1/(1 + exp(-1 * (0))) ## equally rated, playing as black
1/(1 + exp(-1 * (100 * gamma_ests_gm[2])))
```

### Winner and Player Effects

```{r figure 4 right - player level parameters for gm data}
player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_gm_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

## see what the smallest and largest posterior mean is 
## for the winner effects
player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_max(order_by = post_mean, n = 5)

player_pars |> 
  group_by(player_name, param) |> 
  summarise(post_mean = mean(value)) |> 
  filter(param == 2) |> 
  ungroup() |> 
  slice_min(order_by = post_mean, n = 5)

## count how many intervals contain 0
## see how many contain 0
player_pars |> 
  group_by(player_name, param) |> 
  filter(param == 2) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  filter(contain_zero == 0)

## then the winner effects plot
winner_plot_gm <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |>
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Estimated W/L-Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

winner_plot_gm

ggsave(plot = winner_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_prev_winner.png")), dpi = 600,
       height = 7, width = 7)
```

```{r figure s4 right - plot player effects}
#| eval: false
## then repeat for player effects
indiv_plot_gm <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(y = reorder(player_name, value, FUN = median), x = value)) +
  stat_histinterval() +
  labs(x = "Estimated Player Effect", y = element_blank(), 
       title = "GM Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

indiv_plot_gm

ggsave(plot = indiv_plot_gm, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S4_right.png")), dpi = 600,
       height = 7, width = 7)
```

```{r interpret changes in win probability for gm}
#| eval: false
## interpret the individual winner effects
## assume wins 50% of all games
1/(1 + exp(-1 * (0)))
## if won previous
1/(1 + exp(-1 * (0 + 0.216 * 0.5)))
## if lost previous
1/(1 + exp(-1 * (0 + 0.216 * - 0.5)))
## what about if win probability overall was 40%
1/(1 + exp(-1 * (0 + 0.216 * 0.6)))
## 35%, pretty low
1/(1 + exp(-1 * (0 + 0.216 * 0.65)))
```

## Other Comparisons

## GM Prev vs Previous 10

```{r figure 5 left - compare global params n-1 vs n-10}
## global effects first
data_path <- here("results_2025/lichessGrandmasters")
fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. Game")

pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_global <- bind_rows(prev_pars, pre_10_pars) |> 
  ggplot(aes(x = value, color = model)) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_global.png")), dpi = 600,
       height = 3.5, width = 7)
```

```{r}
#| eval: false
## TO DO - move to supplement
## not required for a paper plot
player_pars_pre <- fit_gm_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Prev. Game")

player_pars_pre_10 <- fit_gm_bullet_10$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_winner <- player_pars_pre |> 
  filter(param == 2) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 2) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

hist_comp_winner

ggsave(plot = hist_comp_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_winner.png")), dpi = 600,
       height = 5, width = 7)

## check the proportion of 95% intervals which contain zero
bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(param == 2) |> 
  group_by(player_name, model) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |> 
  group_by(model) |> 
  summarise(prop = sum(contain_zero)/ n())

## do the effects change in direction
bind_rows(player_pars_pre, player_pars_pre_10) |> 
  filter(param == 2) |> 
  group_by(player_name, model) |> 
  summarise(lower = quantile(value, prob = 0.025),
            upper = quantile(value, prob = 0.975)) |> 
  mutate(contain_zero = ifelse(upper > 0 & lower < 0,
                         1, 0)) |>
  print(n = 50)

## then player effects
## should update this also
hist_comp_player <- player_pars_pre |> 
  filter(param == 1) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_pre_10 |> filter(param == 1) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

hist_comp_player
ggsave(plot = hist_comp_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "gm_comp_player.png")), dpi = 600,
       height = 5, width = 7)
```

## Bullet vs Blitz for 17-19

Compare the global parameters and the winner effects here, using the
previous game only.

```{r bullet v blitz for 17-19}
#| eval: false

## Potential Supplement Plots
data_path <- here("results_2025/lichess1700-1900")
fit_17_19_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev_n1.RDS"
))
bullet_17_19 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_17_19_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev_n1.RDS"
))
blitz_17_19 <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))

player_id_17_19 <- as_tibble(bullet_17_19) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

## compare global effects first
bullet_pars <- fit_17_19_bullet_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Bullet")

blitz_pars <- fit_17_19_blitz_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Blitz")

bul_blit_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
  ggplot(aes(value, color = model))  +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

bul_blit_comp_global

ggsave(plot = bul_blit_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_global.png")), dpi = 600,
       height = 5, width = 7)

## then repeat for winner and player effects
player_pars_bullet <- fit_17_19_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id_17_19, by = "id") |> 
  mutate(model = "Bullet")

player_pars_blitz <- fit_17_19_blitz_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id_17_19, by = "id") |> 
  mutate(model = "Blitz")

## need to keep only players which are in both 
## bullet and blitz

common_players <- intersect(unique(player_pars_bullet$player_name),
                            unique(player_pars_blitz$player_name))

comp_17_19_winner <- player_pars_bullet |> 
  filter(param == 2) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 2) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

comp_17_19_winner

ggsave(plot = comp_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

comp_17_19_player <- player_pars_bullet |> 
  filter(param == 1) |> 
  filter(player_name %in% common_players) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |>
              filter(param == 1) |>
              filter(player_name %in% common_players) |>
              mutate(y = as.factor(player_name))) |>
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

comp_17_19_player

ggsave(plot = comp_17_19_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_17_19_player.png")), dpi = 600,
       height = 5, width = 7)
```

## Bullet vs Blitz for GM

```{r figure 5 right - bullet v blitz gm}
data_path <- here("results_2025/lichessGrandmasters")
fit_gm_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev_n1.RDS"
))
bullet_gm <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))
fit_gm_blitz_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_blitz_model_prev_n1.RDS"
))
blitz_gm <- readRDS(file = paste0(data_path, "/users_blitz.RDS"))
## compare global effects first
bullet_pars <- fit_gm_bullet_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Bullet")

blitz_pars <- fit_gm_blitz_prev$draws(c(
  "mu_beta",
  "gamma1", "gamma2"
)) |>
  as_draws_df() |>
  pivot_longer(cols = "mu_beta":"gamma2") |>
  mutate(model = "Blitz")

bul_blit_gm_comp_global <- bind_rows(bullet_pars, blitz_pars) |>
  ggplot(aes(x = value, y = name, color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  facet_wrap(~name,
    scales = "free",
    labeller = labeller(name = facet_labels)
  ) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

bul_blit_gm_comp_global

ggsave(
  plot = bul_blit_gm_comp_global,
  filename = paste0(here(
    "Paper_WriteUp", "paper_figures",
    "bul_blitz_comp_gm_global.png"
  )), dpi = 600,
  height = 3.5, width = 7
)
```

```{r figure 6 - comparison between bullet and blitz winner}
## then repeat for winner and player effects
player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars_bullet <- fit_gm_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Bullet")

player_pars_blitz <- fit_gm_blitz_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  mutate(model = "Blitz")

comp_gm_winner <- player_pars_bullet |> 
  filter(param == 2) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |> 
              filter(param == 2) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  scale_color_viridis_d(option = "H")

comp_gm_winner
ggsave(plot = comp_gm_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_gm_winner.png")), dpi = 600,
       height = 5, width = 7)
```

```{r}
#| eval: false
## player effects also
comp_gm_player <- player_pars_bullet |> 
  filter(param == 1) |> 
  mutate(y = reorder(player_name, value, FUN = median)) |> 
  bind_rows(player_pars_blitz |> 
              filter(param == 1) |> 
              mutate(y = as.factor(player_name))) |> 
  ggplot(aes(x = value, y = y,
             color = model)) +
  stat_pointinterval(position = ggstance::position_dodgev(height = 0.5)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_single_y() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  scale_color_viridis_d(option = "H")

comp_gm_player

ggsave(plot = comp_gm_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "bul_blitz_comp_gm_player.png")), dpi = 600,
       height = 5, width = 7)
```

# Section 5: Model Validation

We currently have two initial model checking 
procedures.

### Permuting for 1700-1900

```{r perm for 17-19}
#| eval: false
perm_data_path <- here("results/lichess1700-1900/perm")

path_id <- 1
data_path <- res_data_path[path_id]
(perm_files <- list.files(perm_data_path))

perm_17_19 <- perm_files[1:1000] |>
  map_dfr(~ read_rds(here(perm_data_path, .x)))

sim_pars <- perm_17_19 |>
  select(variable, mean, sim_id)

fit_17_19_bullet_prev <- readRDS(file = paste0(
  data_path,
  "/all_rated_bullet_model_prev.RDS"
))

bullet_17_19 <- readRDS(file = paste0(
  data_path,
  "/users_bullet.RDS"
))

## then need to overlay the true values also
true_values <- fit_17_19_bullet_prev$summary() |>
  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |>
  select(variable, mean)

perm_17_19_global <- sim_pars |>
  filter(variable %in% c(
    "mu_beta", "gamma1", "gamma2")) |>
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free") +
  geom_vline(
    data = true_values,
    mapping = aes(xintercept = mean),
    col = "red"
  ) +
  theme_3col() +
  scale_x_continuous(breaks = breaks_pretty(n = 3)) +
  labs(x = element_blank(), y = element_blank())

perm_17_19_global

ggsave(
  plot = perm_17_19_global,
  filename = paste0(here(
    "Paper_WriteUp", "paper_figures",
    "perm_17_19_global.png"
  )), dpi = 600,
  height = 5, width = 7
)
```

```{r figure 8 left - perm winner effects}
## get the selected users for this, top 20 players
path_id <- 1  ## for 1700-1900
data_path <- all_data_path[path_id]
files <- list.files(data_path)

lichess_data <- files |> 
  map_dfr(~read_player(data_path, .x))

bullet_data <- lichess_data |> 
  mutate(Event = tolower(Event)) |>
  filter(TimeControl == "60+0") |>
  filter(Variant == "Standard") |>
  filter(grepl("rated bullet game", Event))

select_users <- bullet_data |> 
  group_by(Username) |> 
  tally() |> 
  arrange(-n) |> 
  ## will change this to n = 20 when run on cluster potentially
  slice_max(order = n, n = 20) |> 
  pull(Username)

## then repeat for the winner effects
player_id <- as_tibble(select_users) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- sim_pars |> 
  as_draws_df() |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  drop_na() |> 
  select(param, id, mean, sim_id) |> 
  mutate(id = as.numeric(id)) |> 
  left_join(player_id, by = "id")

perm_players <- unique(player_pars$player_name)

true_beta <- fit_17_19_bullet_prev$summary() |> 
  filter(str_starts(variable, "beta\\[")) |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  select(param, id, mean) |> 
  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  filter(player_name %in% perm_players)

levels <- true_beta |> 
  filter(param == 2) |> 
  arrange(mean) |> 
  pull(player_name)

true_vals <- true_beta |> 
  filter(param == 2) |>  
  mutate(y = row_number() - 0.5,
         yend = y + 1)

perm_17_19_winner <- player_pars |> 
  filter(param == 2) |>
  mutate(y = factor(player_name, levels = levels)) |> 
  ggplot(aes(x = mean, y = y)) +
  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
  geom_segment(data = true_vals, 
               aes(x = mean, y = y, yend = yend), col = "red") +
  labs(y = "", x = "Estimated W/L-Effect") +
  theme_single_y()

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

```

## GMs

### Permuting for GM Data

```{r figure 8 right}
perm_data_path <- here("results/lichessGrandmasters/perm/")

## read all the files in this directory
perm_files <- list.files(perm_data_path)

perm_gm <- perm_files[1:1000] |> 
  map_dfr(~read_rds(here(perm_data_path, .x)))

sim_pars <- perm_gm |> 
  select(variable, mean, sim_id)

## then need to overlay the true values also
true_values <- fit_gm_bullet_prev$summary() |> 
  filter(variable %in% c("mu_beta", "gamma1", "gamma2")) |> 
  select(variable, mean)

perm_gm_global <- sim_pars |> 
  filter(variable %in% c("mu_beta",  "gamma1", "gamma2",
                           "sigma_1", "tau[1]", "tau[2]",
                           "sigma_g1", "sigma_g2")) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free") +
  geom_vline(data = true_values, mapping = aes(xintercept = mean),
             col = "red") +
  theme_3col() +
  labs(x = element_blank(), y = element_blank())

ggsave(plot = perm_gm_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_global.png")), dpi = 600,
       height = 5, width = 7)

## then repeat for the winner effects
player_id <- as_tibble(bullet_gm) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- sim_pars |> 
  as_draws_df() |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  drop_na() |> 
  select(param, id, mean, sim_id) |> 
  mutate(id = as.numeric(id)) |> 
  left_join(player_id, by = "id")

perm_players <- unique(player_pars$player_name)

true_beta <- fit_gm_bullet_prev$summary() |> 
  filter(str_starts(variable, "beta\\[")) |> 
  mutate(param = stringr::str_extract(variable, pattern = "\\d"),
         id = stringr::str_extract(variable, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  select(param, id, mean) |> 
  mutate(id = as.numeric(id), param = as.numeric(param)) |> 
  left_join(player_id, by = "id") |> 
  filter(player_name %in% perm_players)

levels <- true_beta |> 
  filter(param == 2) |> 
  arrange(mean) |> 
  pull(player_name)

true_vals <- true_beta |> 
  filter(param == 2) |>  
  mutate(y = row_number() - 0.5,
         yend = y + 1)

perm_gm_winner <- player_pars |> 
  filter(param == 2) |>
  mutate(y = factor(player_name, levels = levels)) |> 
  ggplot(aes(x = mean, y = y)) +
  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
  geom_segment(data = true_vals, 
               aes(x = mean, y = y, yend = yend), col = "red") +
  theme_single_y() +
  labs(x = "Estimated W/L-Effect", y = "")

perm_gm_winner

ggsave(plot = perm_gm_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_winner.png")), dpi = 600,
       height = 5, width = 7)

levels <- true_beta |> 
  filter(param == 1) |> 
  arrange(mean) |> 
  pull(player_name)

true_vals <- true_beta |> 
  filter(param == 1) |>  
  mutate(y = row_number() - 0.5,
         yend = y + 1)

perm_gm_player <- player_pars |> 
  filter(param == 1) |>
  mutate(y = factor(player_name, levels = levels)) |> 
  ggplot(aes(x = mean, y = y)) +
  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
  geom_segment(data = true_vals, 
               aes(x = mean, y = y, yend = yend), col = "red") +
  theme_single_y() +
  labs(x = "Estimated Player Effect", y = "")

perm_gm_player

ggsave(plot = perm_gm_player, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_gm_player.png")), dpi = 600,
       height = 5, width = 7)

```

<!-- TO DO - Tidy up below here -->

## Posterior Predictive Plots

```{r}

```

# Supplementary Material

This contains all analysis and figures which are included in 
the supplementary materials.

## Additional Summary Information

```{r}
#| eval: false

## how many times does a focal player play another focal player

focal_17_19 <- unique(bullet_data$Username)

bullet_data |> 
  mutate(opponent = ifelse(Username == White, Black, White)) |> 
  mutate(focal_opp = ifelse(opponent %in% focal_17_19, 1, 0)) |> 
  select(Username, focal_opp) |> 
  summarise(sum(focal_opp)/ n())

focal_gm <- unique(bullet_data_gm$Username)

bullet_data_gm |> 
  mutate(opponent = ifelse(Username == White, Black, White)) |> 
  mutate(focal_opp = ifelse(opponent %in% focal_gm, 1, 0)) |> 
  select(Username, focal_opp) |> 
  summarise(sum(focal_opp)/ n())

```

## Results for Other Cohorts

## 2000-2200 Previous Game Winner/Loser Effect

```{r figure s1-left}
#| eval: false
data_path <- here("results_2025/lichess2000-2200")

fit_20_22_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

bullet_20_22 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

player_id <- as_tibble(bullet_20_22) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_20_22_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

winner_plot_20_22 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "2000-2200 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = winner_plot_20_22, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S1_left.png")), dpi = 600,
       height = 7, width = 7)
```

## 2300-2500 Previous Game Winner/Loser Effect

```{r figure-s1-right}
#| eval: false
data_path <- here("results_2025/lichess2300-2500")

fit_23_25_bullet_prev <- readRDS(file = 
                                   paste0(data_path,
                                          "/all_rated_bullet_model_prev_n1.RDS"))

bullet_23_25 <- readRDS(file = paste0(data_path, "/users_bullet.RDS"))

player_id <- as_tibble(bullet_23_25) |> 
  mutate(id = row_number()) |> 
  rename(player_name = value)

player_pars <- fit_23_25_bullet_prev$draws() |> 
  as_draws_df() |>
  select(starts_with("beta[")) |>
  pivot_longer(cols = everything()) |>
  mutate(param = stringr::str_extract(name, pattern = "\\d"),
         id = stringr::str_extract(name, pattern = "\\d+]"),
         id = stringr::str_replace(id, "\\]", ""),
         player_id = paste0("beta[", id, "]")) |> 
  mutate(id = as.numeric(id),
         param = as.numeric(param)) |> 
  left_join(player_id, by = "id")

winner_plot_23_25 <- player_pars |> 
  filter(param == 2) |> 
  mutate(prob = 1/(1 + exp( -1 * value)) - 0.5) |> 
  ggplot(aes(y = reorder(player_name, prob, FUN = median), x = prob)) +
  stat_histinterval() +
  labs(x = "Change in Win Probability", y = element_blank(), 
       title = "2300-2500 Cohort") +
  geom_vline(aes(xintercept = 0), col = "red", linetype = "dashed") +
  theme_single_y()

ggsave(plot = winner_plot_23_25, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S1_right.png")), dpi = 600,
       height = 7, width = 7)
```

## Fitting the Model with n=5 as History

To be added here for 1700-1900 and GM cohorts.

```{r figure S3 left}
## global effects first
data_path <- here("results_2025/lichess1700-1900")

fit_17_19_bullet_5 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))

fit_17_19_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

prev_pars <- fit_17_19_bullet_prev$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. Game")

pre_5_pars <- fit_17_19_bullet_5$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 5 Games")

pre_10_pars <- fit_17_19_bullet_10$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
  ggplot(aes(x = value, color = model)) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S3_left.png")), dpi = 600,
       height = 3.5, width = 7)

```

```{r figure S3 right}
## global effects first
data_path <- here("results_2025/lichessGrandmasters")

fit_gm_bullet_5 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n5.RDS"))

fit_gm_bullet_10 <- 
  readRDS(file =paste0(data_path,"/all_rated_bullet_model_n10.RDS"))

prev_pars <- fit_gm_bullet_prev$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. Game")

pre_5_pars <- fit_gm_bullet_5$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 5 Games")

pre_10_pars <- fit_gm_bullet_10$draws(c("mu_beta",
                                        "gamma1", "gamma2")) |> 
  as_draws_df() |> 
  pivot_longer(cols = "mu_beta":"gamma2") |> 
  mutate(model = "Prev. 10 Games")

hist_comp_global <- bind_rows(prev_pars, pre_5_pars, pre_10_pars) |> 
  ggplot(aes(x = value, color = model)) +
  stat_pointinterval(position = position_dodge(width = 0.2)) +
  facet_wrap(~name,
             scales = "free",
             labeller = labeller(name = facet_labels)) +
  labs(x = element_blank(), y = element_blank(), color = element_blank()) +
  theme_3col() +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = breaks_pretty(n = 2)) +
  scale_color_viridis_d(option = "H")

ggsave(plot = hist_comp_global, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "Figure_S3_right.png")), dpi = 600,
       height = 3.5, width = 7)

```